### 2024-06-17 README.Rmd
Working on corporate side of things

Notice that this statement (like the definition of the work before the panels above), takes for granted that 1) AIs are powerful tools, and 2) universities can be leading players in their adoption. This attitude is almost fatalistic^[In the sense of "fated to".] in its enthusiasm, and perfectly consistent with the hype observed around AI this time^[The pattern of cyclical hypes around AI even has a name: [AI winters](https://en.wikipedia.org/wiki/AI_winter); the current one, with chatbots playing the loudest horn of the bandwagon, is just the most recent of several cycles of hype [see @siegel_ai_2023].] [@angwin_press_2024].

@@

We also need to understand the goals of the organization that considers using the technology. When does it make sense for a university like McMaster to "integrate" AIs as part of the core mission of the organization, accounting for what the technology does for some people and what it does to others?

@@

@@

A key reason why many organizations have decided to spend substantial resources considering the adoption of AIs is that, unlike most other tools, they have been presented quite deliberately as _everything tools_^[Almost as flexible as, well, a naturally intelligent human, but faster and supposedly cheaper.]. While hammers are understood to have a limited range of uses that rely on concentrating force, and word processors are mostly used to record words, the release of GenAIs to the masses was accompanied by innumerable lofty promises^[Quite a few of them made by entities and individuals that had a stake in the adoption of the technology, ranging from "AI is inevitable, and I will sell it to you" to "AI is inevitable, and for a consulting fee I will tell you how to level up with it".].

@@

According to Allerin^[Curiously, Allerin is not identified as an artificial intelligence company in @joshi_5_2022 but only in the much later article by @garsten_what_2024.], AI-powered self-driving cars would reduce accidents, car ownership, pollution and noise, and would automate logistics and make parking tickets obsolete [@joshi_5_2022; @garsten_what_2024]. But there is no reason to stop at self-driving vehicles: McMaster University's Canada Excellence Research Chair Laureate Ali Emadi "hopes to help revolutionize and influence everything from personal travel to city planning" by putting AI directly into our transportation systems [@dorey_driving_2023].

@@

AI will even write code, perhaps helping itself to some self-programming along the way: AIs that write themselves, or AI self-help?

@@

The forecast that AI will improve everything (from the natural environment of the globe to the dumps of our cats) was certain to, sooner or later, reach higher-education, a target-rich environment for selling hype--and indeed, colleges and universities have not been spared the onslaught of promises. Examples of this abound.

@power_3_2024,^[Rhett Power is a motivational speaker and executive coach.], for one, presents three innovative use cases for AI in higher education. First, AI could take over manual tasks and improve the budget by releasing administrative staff^[Probably to work on something else, and not for the university, hence leading to "efficiencies".]. AI could evaluate applications for admissions, including assessing traits like the "grit" and "empathy" of applicants. AI could also personalize the experience of consumers (occasionally also called students) by predicting their final grade with 60-70\% accuracy, and by providing corrective and instructive feedback.

@@

In another instance, @hie_how_2023^[Anthony Hié is Chief Innovation and Digital officer at [Excelia](https://www.excelia-group.com/about-excelia/our-schools/excelia-business-school). Claire Thouary is the founder of [QACE UP](http://qace-up.fr), a consulting firm that supports higher-education institutions in their quest for "quality excellence", whatever that means.] argue that AIs will likely be the primary way to access information in the future. For this reason, institutions must equip students with the skills needed to use AI competently, in particular train them to perfect the art of _prompt engineering_ (the ability to craft effective queries that prompt a GenAI to provide the most useful outputs). Universities will use AIs to improve learning and challenge the critical thinking of students. Faculty will also benefit from AI: it will be used to evaluate their students^[A task of dubious legality, and clearly a way to cheat students of a learning experience.], and will handle repetitive tasks like writing course syllabi^[In my 20+ years of experience as an educator, writing syllabi has been essential to understanding my own courses, what I teach and how I teach it; automating the task of writing syllabi is another way of degrading the quality of instruction that students can expect to receive.].

@@

A benefit of portraying AI as an _everything tool_ (for the proponents of the technology, at least), is that conversations around it quickly turn into a [Gish Gallop](https://en.wikipedia.org/wiki/Gish_gallop), a rhetorical device where a party makes an overwhelming number of arguments with little regard to their accuracy or strength. To be sure, AIs _can_ do some things well, but critics are forced to respond to every single propounded use case with in-depth analysis, which usually is much more time consuming than making the unsubstantiated and/or misleading claims of the gallop itself^[This communication strategy is what Trump's erstwhile and possibly future advisor Steve Bannon termed ["flooding the zone with shit"](https://www.vox.com/policy-and-politics/2020/1/16/20991816/impeachment-trial-trump-bannon-misinformation). Perhaps not coincidentally, AI is now a major vector of disinformation, flooding the zone at a rate that humans cannot hope to counteract [@maiberg__google_2024].]. Proponents of AI can always admit that, ok, AIs do not actually do _that one thing_, but what about the myriad of other things that people say they really do do?

@@

In what follows, it is not my intention to confront every single possible use case of AI in teaching and learning, since doing so plays into the intent of the Gish Gallop^[I have already spent more than 75 hours working on this essay, and it is still nowhere in the vicinity of being an exhaustive apology or indictment of AI.], namely to defeat criticism by flooding the zone. As well, I reject the dilemma posed by the purveyors of hype, and I remain convinced that there are multiple ways to excel at teaching and learning that do not require AI, let alone the moral, environmental, and social compromises demanded by AI^[Most discussions of AI as an everything tool acknowledge that the use of AIs is not without problems. For instance, @mollick_instructors_2024 note that "Large Language Models are trained in ways that may violate copyright, and often rely on the efforts of low-wage workers in precarious conditions for part of their training. Models are trained on biased data and can produce either subtly or overtly biased results. And because these biases seem to come from an objective machine, they can be especially pernicious. Using AI systems can mean sharing private data with the for-profit companies developing LLMs, and that data may be used to train further AIs." Therefore, "educators may want to consider before deciding whether to use these systems". This disclaimer, all one third of a page in a 76-page document dedicated to extolling an AI-enabled future in education, highlights the important role of institutions to provide the moral and ethical clarity that may elude individual instructors working under desperately under-resourced conditions, or simply willing to win a race to the bottom.].

@@

Below is an example of the simplest regression model with only one independent variable (i.e., a bivariate model). The plot shows the pairs of $y$ and $x$ values, and the line is the model, the conditional mean $\hat{y}$. This model needs only two parameters: an intercept (the conditianl mean $\hat{y}$ when $x = 0$), and the slope of the line, in other words, the rate of change of $\hat{y}$ with respect to $x$. Each additional independent variable uses one additional parameter to describe the slope of the model with respect to that variable.
```{r simulate-simple-example}
set.seed(13253)

@@

Indeed, the first neural network is "better" than the linear regression in that it reduces the typical error by `r paste0(prettyNum(delta_mad_nn1, digits = 2), "%")`, whereas the deeper network reduces the typical error by `r paste0(prettyNum(delta_mad_nn2, digits = 2), "%")`, also with respect to the linear regression^[The "typical error" here is the mean absolute deviation of the predictions.].

There is a limit as to how flexible a model can be, and this is true of linear regression and of neural networks. Each additional parameter makes the model more flexible, but as a rule we cannot use more parameters than the number of data points that exist in the dataset used to train the model. Roughly, every additional parameter needs a data input to train on, and when there are as many parameters as data inputs the model becomes overfit and will predict each value of the independent variable in the training dataset perfectly. The model will fail to train if there are more parameters than data points.

@@

- AIs are neural networks, which is to say, regression models.
- A regression model generates outputs that are regressions to the mean (i.e., statistically probable outputs).
- AIs are _very_ flexible non-linear regression models.
- Their flexibility comes from using "deep" architectures, which means many layers with many, many parameters^[According to reports [see @griffith_gpt_4_2023], GPT-1 had 117 million parameters, GPT-2 1.5 billion, and GPT-3 175 billion parameters. OpenAI no longer reports the number of parameters of its newer models, but in the latest version this is estimated to be in the trillions.]. The large number of parameters makes these models _extremely_ data hungry. 
- AIs are trained using an initial set of random values as parameters. The same training dataset can result in different models due to the randomness in the initial conditions.
- The flexibility of AIs means that outputs will vary seemingly at random with relatively small changes in the prompt. 
- Also, importantly, since regression models aim for the mean (i.e., a statistically probable output), they tend to be bad at _extrapolation_ (or thinking outside the box, if AIs could think). The further away an input (i.e., prompt) is from the space defined by the training dataset, the worse the output will tend to be^[Which is possibly why so-called hallucinations became less common as more and more data were used to train AIs. The increasingly large datasets essentially expand the space where the model could interpolate.].

@@

Next, I lay out some considerations about the use of AI for teaching and learning, in an attempt to find out where it makes sense to use this technology. To begin, we must note the number of examples of the use of AI by students. A Turnitin-sponsored  report [@tyton_partners_generative_2023] of 1,600 post secondary students and 1,000 faculty, found that 50% of students used GenAI, compared to only 22% faculty. Use of AI by students is commonly cited as a reason for universities to embrace AI [e.g., @power_3_2024]. McMaster's own internal research (based on a sample of over 1,300 students) indicates that at least 61% of students used AI during the Fall Term in 2023, 75% of them for school work [@yachouh_student_2024].

@@

Most textbooks already include plenty of practice questions, especially for relatively common topics. For more specialized topics, asking interesting questions should be a skill in and of itself. An emphasis on practice questions, however, signals an underlying weakness of the process, namely, a focus on learning for the examination, as opposed to learning for its own sake. A growing body of evidence points at the effectiveness of alternative assessment schemes, such as continuous assessment through course work, and even seen examinations instead of the traditional unseen examinations [@turner_see_2018]. Alas, unseen examinations continue to dominate the evaluation landscape [@buckley_are_2023]. Ungrading, in particular, is increasingly recognized for its potential to help learners internalize the motivation to learn, and for fostering adaptability, creative thinking, and self-management [@gorichanaz_it_2024].

@@

There is nothing new about automated grading: [Scantron](https://www.scantron.com/company/) sheets for multiple choice examinations have existed for over four decades. They work well because the algorithm needed for grading multiple choices is simple, and importantly, it is deterministic (i.e., there is no randomness involved). There is no reason to use a neural network to grade multiple choice examinations. But on the flip side, there is also no evidence that a neural network can perform well at grading anything more complex than a multiple choice examination, especially since 1) it will tend to consider things that are closer to the mean as "better"^[My own unpublished experiments with grading written reflections were underwhelming, missing the human-assigned grade over 30\% of the time]; and 2) the amount of training required for a classification task of this nature is massive^[Labeling data is an essential, yet extremely expensive aspect of training GenAIs. The algorithm needs a "dependent" variable for training, in this case a grade assigned to a piece of work: this is called a label. Grading assignments manually is the process of labeling. Given all the possible variations of assignments (by year, by subject, by instructor), a general grading machine would require vast volumes of _labelled_ training data, and even so there is no guarantee that the results will be adequate, let alone appropriate.].

@@

An AI should not be used to identify signs of learning difficulties for the same reasons that AIs cannot be trusted to identify text written by an AI. The false positive rates (blaming the innocent) are out of proportion compared to the benefits the practice might offer. A vigilant instructor will be aware of learning difficulties among their students (but not if the students are an ocean of anonymous faces). But were the process to become automated, who would decide that an intervention is needed, and who qualified for it? Would the intervention be offered to everyone flagged by the system, including the false positives? And, could those denied an intervention complain? Importantly, who will fund any intervention? And who will deliver it? Another AI?

@@

Reliance on AI can short-circuit the learning process. Statements like "in the very near future AIs will likely be the primary way we access knowledge" [@hie_how_2023] convey the idea that prompt engineering will become an essential skill; perhaps even the skill to end all skills^[Prompt engineering is a pompous name for a task that amounts to trying to read the entrails of a Large Language Model: something that cannot possibly be performed in a systematic way, and that offers no guarantees that it will work in the same way again, especially after a model is updated. No genuine engineer would do their work in such fashion, and calling the activity of writing prompts "engineering" is a blatant attempt to steal the competence and reputation of actual engineering.].

It is important to note that, even if prompt engineering offered some value to students, the skill is inextricable tied to a singular technology, with little transferability potential. As @mollick_instructors_2024 demonstrate, prompt engineering is not even transferable _between_ different LLMs (see p. 5 and p. 36). Good skills transfer between contexts. Learning how to form an effective bibliographical query would work at retrieving relevant resources at McMaster, at the University of Toronto, and at the New York Public Library. Learning to identify relevant content would work while reading a novel, a book of history, or a text of econometrics. Writing, the slow and painful process of converting thoughts into words, is essential to create a style of communication that is unique to every person. It might be argued, in fact, that writing _is_ learning [@klein_trends_2016].

In a sense, an AI is like a crutch. It can help someone to accomplish a task more easily, but it will not help if used constantly. However, what the public is being promised is not a crutch: it is an All-Terrain Vehicle that the end user cannot hope to master^[Reports of people getting their fingers crushed by Tesla's cybertruck could be a result of 1) people trusting the technology more than they should; and 2) not understanding the technology: apparently the algorithm uses increasing force to close the trunk if it fails to latch on back-to-back cycles. Hence, the trunk will not slash a banana on its first attempt to latch, but will crush a finger on the second attempt [@lambert_tesla_2024].] due to its blackbox nature--in the sense that 1) neural networks are blackboxes; and 2) the development of AIs itself is a blackbox where things change based on the interests of their corporate developers^[Sometimes said interest may be relatively benign, like avoiding reputational damage [by correcting a wrong](https://www.bbc.com/news/technology-65202597), but there is no guarantee that corporate interests will always align with those of the public, and in fact, they seldom will as long as there are only a handful of American companies that offer an "essential" service.] instead of the interests of their relatively small customers^[And this refers not only to individuals whose skills are yoked to ChatGPT and MidJourney. In 2022, McMaster University operated on revenues of less than [\$843 million](https://financial-affairs.mcmaster.ca/app/uploads/2023/06/2023-24-Consolidated-Budget-FINAL.pdf) (CAD). Microsoft, in contrast, operated on revenues of approximately [\$198,000 millions](https://www.microsoft.com/investor/reports/ar22/) (USD) in 2022. To put things in perspective (and using the average exchange rate of 2022), if McMaster used _all_ of its 2022 revenue to pay Microsoft, that would leave McMaster with zero dollars and still account for less than _one third of one percent point_ of Microsoft's revenue. A mid-sized Canadian university is like an ant scuttling under the economic boot of Microsoft. In truth, Microsoft feels that it can stand up to even the US Government. As seen in the case of Solar Winds, Microsoft refused to fix a known security flaw that left the US Government vulnerable to a Russian hack, and did so with little fear of ever facing accountability [@burke_microsoft_2024].].

@@

But let us assume for a moment that using AIs in their current form does becomes an in-demand tool in the short term. What is the expected longevity of skills like prompt engineering?

Apart from a few methodological developments (like transformers), improvement in the performance of AIs (say, from ChatGPT-2 to the latest version), mostly are the result of using bigger datasets. This has led to a rush, as developers try to improve their models by training them on more and more data^[A rush that feels at least as frenetic as any gold rush in the past. And, while the gold rushes of the 19th century were relatively democratic due to their low entry costs--both in terms of immigration and tools--the current data rush has enormous entry costs and is largely dominated by Venture Capital (VC) and large corporations.]. But, despite the large number of text items and images available in digital form, there is still only a finite amount of data to be had.

As of this writing, the few big players that can afford to grab the data, are mostly done with all the data grabbing, sometimes by morally dubious, and at times by possibly illegal means [@clark_ex-amazon_2024]. IBM, for example, uses data drawn from the Project CodeNet Dataset. The data are "sourced" from two online judge web sites, AIZU Online Judge and AtCoder [@ibm_automatic_2021]. According to IBM "[the] end goal of CodeNet is to enable developers to create systems that can modernize existing codebases, as well as fix errors and security vulnerabilities in code". When we inspect the Terms of Use of AIZU (a Japanese academic project), "Commercial uses of the registered codes are prohibited". IBM studiously avoids the word "commercial" in its posts about LLM, foundational models, and CodeNet. One might ask if IBM is doing all this research as a non-commercial effort? Will those developers creating systems, modernizing databases, fixing errors, etc., be able to access this for free, and use the systems for education and research only? What does "commercial uses" even mean?

@@

OpenAI, as a leader in data grabbing, turned to Stack Overflow for more data [@stack_overflow_stack_2024]. Stack Overflow [bills itself](https://stackoverflow.com) as "[a] community-based space to find and contribute answers to technical challenges, and one of the most popular websites in the world". Said community was [unhappy](https://favtutor.com/articles/stack-overflow-community-not-happy-openai/) (to say the least) with the way Stack Overflow dealt with OpenAI, and has come to believe that the terms were abusive: while Stack Overflow's terms of service state that posters cannot revoke permission to use, the site also operates under a Creative Commons license that requires attribution, something that AIs notoriously _do not_ do. Some members of the community have protested the deal with OpenAI by trying to withdraw or change their posts, in what Ars Technica unhelpfully termed [sabotage](https://arstechnica.com/information-technology/2024/05/stack-overflow-users-sabotage-their-posts-after-openai-deal/). It is an open question how long Stack Overflow will remain a community-based space, after it started treating its members as unwitting, and even unwilling providers of data for OpenAI.

@@

The implication should be clear: training models with the output of models is a recipe for less varied outputs. We may be witnessing _Peak ChatGPT_, and future versions of AIs, instead of improving, may begin to regress to a bland version of their own previous outputs. And this is before we even consider the possibility of AIs being deliberately "poisoned". As backlash against a data rush that is perceived as exploitative and abusive grows, new tools are being developed to generate data that are manipulated in ways, too subtle for the changes to be detectable by a human, but deleterious to the training process of a GenAI [@salman_raising_2023]. One such tool is @shan_nightshade_2024's _Nightshade_, an algorithm that works to corrupt prompts used in text-to-image generation, and that notably bleeds through to poison related concepts. Another tool is @xu_shadowcast_2024's _Shadowcast_, an algorithm that poisons samples to trick an AI into misidentifying class labels, and that also can perform persuasion attacks, leveraging image-to-text generation capabilities to craft nonsensical narratives.

>>>>

Notice that this statement (like the definition of the work before the panels above), takes for granted that 1) AIs are powerful tools, and 2) universities can be leading players in their adoption. This attitude is almost fatalistic^[In the sense of "fated to".] in its enthusiasm, and perfectly consistent with the hype observed around AI this time^[The pattern of cyclical hypes around AI even has a name: [AI winters](https://en.wikipedia.org/wiki/AI_winter); the current one, with chatbots playing the loudest horn atop the bandwagon, is just the most recent of several cycles of hype [see @siegel_ai_2023]--fundamentally a vastly overblown tech bubble [@vinsel_dont_2023].] [@angwin_press_2024].

@@

We also need to understand the goals of the organization that considers using the technology. When does it make sense for a university like McMaster to "integrate" AIs as part of its core mission, accounting for what the technology does for some people and what it does to others?

@@

And so, what do AIs do that can advance the mission of the university?

@@

A key reason why many organizations have decided to spend substantial resources considering the adoption of AIs is that, unlike most other tools, they have been presented quite deliberately as _everything tools_^[Almost as flexible as, well, a naturally intelligent human, but faster and supposedly cheaper.]. While hammers are understood to have a limited range of uses that mostly have to do with concentrating force, and word processors are generally only used to record words, the release of GenAIs to the masses was accompanied by innumerable lofty promises^[Quite a few of them made by entities and individuals that had a stake in the adoption of the technology, ranging from "AI is inevitable, and I will sell it to you" to "AI is inevitable, and for a consulting fee I will tell you how to level up with it".].

@@

According to Allerin^[Curiously, Allerin is not identified as an artificial intelligence company in @joshi_5_2022 but only in the much later article by @garsten_what_2024.], AI-powered self-driving cars would reduce accidents, car ownership, pollution and noise, and would automate logistics and make parking tickets obsolete [@joshi_5_2022; @garsten_what_2024]. But why stop at self-driving vehicles? McMaster University's Canada Excellence Research Chair Laureate Ali Emadi "hopes to help revolutionize and influence everything from personal travel to city planning" by putting AI directly into our transportation systems [@dorey_driving_2023].

@@

AIs will even write code, perhaps helping themselves to some self-programming along the way: AIs that write themselves, or AI self-help?

@@

The forecast that AI will improve everything (from global climate to the dumps of our cats) was certain to, sooner or later, reach higher-education, a target-rich environment for selling hype--and indeed, colleges and universities have not been spared the onslaught of promises.

Examples of this abound.

@power_3_2024,^[Rhett Power is a motivational speaker and executive coach.], writing for Forbes, presents three innovative use cases for AI in higher education. First, AI could take over manual tasks and improve the budget by releasing administrative staff^[Probably to work on something else, and not for the university, hence leading to "efficiencies".]. AI could evaluate applications for admissions, including assessing traits like the "grit" and "empathy" of applicants. AI could also personalize the experience of consumers (occasionally also called students) by predicting their final grade with 60-70\% accuracy, and by providing corrective and instructive feedback.

@@

In another instance, @hie_how_2023^[Anthony Hié is Chief Innovation and Digital officer at [Excelia](https://www.excelia-group.com/about-excelia/our-schools/excelia-business-school), a top business school in France. Claire Thouary is the founder of [QACE UP](http://qace-up.fr), a consulting firm that supports higher-education institutions in their quest for "quality excellence", whatever that means.] argue that AIs will likely be the primary way to access information in the future. For this reason, institutions must equip students with the skills needed to use AI competently, in particular train them to perfect the art of _prompt engineering_ (the ability to craft effective queries that prompt a GenAI to provide the most useful outputs). Universities will use AIs to improve learning and challenge the critical thinking of students. Faculty will also benefit from AI: it will be used to evaluate their students^[A task of dubious legality, and clearly a way to cheat students of a learning experience.], and will handle repetitive tasks like writing course syllabi^[In my 20+ years of experience as an educator, writing syllabi has been an activity essential to understand my own courses, what I teach and how I teach it; automating the task of writing syllabi is another way of degrading the quality of instruction that students can expect to receive.].

@@

A benefit of portraying AI as an _everything tool_ (for the proponents of the technology, at least), is that conversations around it quickly turn into a [Gish Gallop](https://en.wikipedia.org/wiki/Gish_gallop), a rhetorical device where a party makes an overwhelming number of arguments with little regard to their accuracy or strength. To be sure, AIs can do _some_ things well, but critics are forced to respond to every single propounded use case with in-depth analysis, which usually is much more time consuming than making the unsubstantiated and/or misleading claims of the gallop itself^[This communication strategy is what Trump's erstwhile and possibly future advisor Steve Bannon termed ["flooding the zone with shit"](https://www.vox.com/policy-and-politics/2020/1/16/20991816/impeachment-trial-trump-bannon-misinformation). Perhaps not coincidentally, AI is now a major vector of disinformation, flooding the zone at a rate that humans cannot hope to counteract [@maiberg__google_2024].]. Proponents of AI can always admit that, ok, AIs do not actually do _that one thing_, but what about the myriad of other things that people say they really do do?

@@

In what follows, it is not my intention to confront every single possible use case of AI in teaching and learning, since doing so plays into the intent of the Gish Gallop^[I have already spent more than 75 hours working on this essay, and it is still nowhere in the vicinity of being an exhaustive apology or indictment of AI.], namely to defeat criticism by flooding the zone. As well, I reject the dilemma posed by the purveyors of hype, and I remain convinced that there are multiple ways to excel at teaching and learning that do not require AI, let alone the moral, environmental, and social compromises demanded by AI^[Most discussions of AI as an everything tool acknowledge that the use of AIs is not without problems. For instance, @mollick_instructors_2024 note that "Large Language Models are trained in ways that may violate copyright, and often rely on the efforts of low-wage workers in precarious conditions for part of their training. Models are trained on biased data and can produce either subtly or overtly biased results. And because these biases seem to come from an objective machine, they can be especially pernicious. Using AI systems can mean sharing private data with the for-profit companies developing LLMs, and that data may be used to train further AIs." Therefore, "educators may want to consider before deciding whether to use these systems". This disclaimer, consuming fully one third of a page in a 76-page document dedicated to extolling an AI-enabled future in education, highlights the important role of institutions to provide the moral and ethical clarity that may elude individual instructors working under desperately under-resourced conditions, or simply willing to win a race to the bottom.].

@@

Below is an example of the simplest regression model with only one independent variable (i.e., a bivariate model). The plot shows the pairs of $y$ and $x$ values, and the line is the model, i.e., the conditional mean $\hat{y}$. This model needs only two parameters: an intercept (the conditianl mean $\hat{y}$ when $x = 0$), and the slope of the line, in other words, the rate of change of $\hat{y}$ with respect to $x$. Each additional independent variable uses one additional parameter to describe the slope of the model with respect to that variable.
```{r simulate-simple-example}
set.seed(13253)

@@

Indeed, the first neural network is "better" than the linear regression in that it reduces the typical error by `r paste0(prettyNum(abs(delta_mad_nn1), digits = 2), "%")`, whereas the deeper network reduces the typical error by `r paste0(prettyNum(abs(delta_mad_nn2), digits = 2), "%")`, also with respect to the linear regression^[The "typical error" here is the mean absolute deviation of the predictions.].

There is a limit as to how flexible a model can be, and this is true of linear regression and of neural networks. Each additional parameter makes the model more flexible, but as a rule we cannot use more parameters than the number of data points that exist in the dataset used to train the model. Roughly, every additional parameter needs one additional datum to train on, and when there are as many parameters as data inputs the model becomes overfit and will predict each value of the independent variable in the training dataset perfectly. The model will fail to train if there are more parameters than data points.

@@

- AIs are neural networks, which is to say, regression models.
- A regression model generates outputs that are regressions to the mean (i.e., statistically probable outputs).
- AIs are _very_ flexible non-linear regression models.
- Their flexibility comes from using "deep" architectures, which means many layers with many, many parameters^[According to reports [see @griffith_gpt_4_2023], GPT-1 had 117 million parameters, GPT-2 1.5 billion, and GPT-3 175 billion parameters. OpenAI no longer reports the number of parameters of its newer models, but in the latest version this is estimated to be in the trillions.]. The large number of parameters makes these models _extremely_ data hungry. 
- AIs are trained using an initial set of random values as parameters. The same training dataset can result in different models due to the randomness in the initial conditions.
- The flexibility of AIs means that outputs will vary seemingly at random with relatively small changes in the prompt. 
- Also, importantly, since regression models aim for the mean (i.e., a statistically probable output), they tend to be bad at _extrapolation_ (or thinking outside the box, if AIs could think). The further away an input (i.e., prompt) is from the space defined by the training dataset, the worse the output will tend to be^[Which is possibly why so-called hallucinations became less common as more and more data were used to train AIs. The increasingly large datasets essentially expand the space where the model can interpolate.].

@@

Next, I lay out some considerations about the use of AI for teaching and learning, in an attempt to find out where it might make sense to use this technology. To begin, we must note the number of examples of the use of AI by students. A Turnitin-sponsored  report [@tyton_partners_generative_2023] of 1,600 post secondary students and 1,000 faculty, found that 50% of students used GenAI, compared to only 22% faculty. Use of AI by students is commonly cited as a reason for universities to embrace AI [e.g., @power_3_2024]. McMaster's own internal research (based on a sample of over 1,300 students) indicates that at least 61% of students used AI during the Fall Term in 2023, 75% of them for school work [@yachouh_student_2024].

@@

Most textbooks already include plenty of practice questions, especially for relatively common topics. For more specialized topics, asking interesting questions should be a skill in and of itself. An emphasis on practice questions, however, signals an underlying weakness of the process, namely, a focus on learning for the examination, as opposed to learning for its own sake. A growing body of evidence points towards the effectiveness of alternative assessment schemes, such as continuous assessment through course work, and even seen examinations instead of the traditional unseen examinations [@turner_see_2018]. Alas, unseen examinations continue to dominate the evaluation landscape [@buckley_are_2023]. Ungrading, in particular, is increasingly recognized for its potential to help learners internalize the motivation to learn, and for fostering adaptability, creative thinking, and self-management [@gorichanaz_it_2024].

@@

There is nothing new about automated grading: [Scantron](https://www.scantron.com/company/) sheets for multiple choice examinations have existed for over four decades. They work well because the algorithm needed for grading multiple choices is simple, and importantly, deterministic (i.e., there is no randomness involved). There is no reason to use a neural network to grade multiple choice examinations. But on the flip side, there is also no evidence that a neural network can perform well at grading anything more complex than a multiple choice examination, especially since 1) it will tend to consider things that are closer to the mean as "better"^[My own unpublished experiments with grading written reflections were underwhelming, missing the human-assigned grade over 30\% of the time]; and 2) the amount of training required for a classification task of this nature is massive^[Labeling data is an essential, yet extremely expensive aspect of training GenAIs. The algorithm needs a "dependent" variable for training, in this case a grade assigned to a piece of work: this is called a label. Grading assignments manually is the process of labeling. Given all the possible variations of assignments (by year, by subject, by instructor), a general grading machine would require vast volumes of _labelled_ training data, and even so there is no guarantee that the results would be adequate, let alone appropriate.].

@@

An AI should not be used to identify signs of learning difficulties for the same reasons that AIs cannot be trusted to identify text written by an AI. The false positive rates (blaming the innocent) are out of proportion compared to the benefits the practice might offer. A vigilant instructor will be aware of learning difficulties among their students (but not if the students are an ocean of anonymous faces). But were the process to become automated, who would decide when an intervention was warranted, and who qualified for it? Would the intervention be offered to everyone flagged by the system, including the false positives? And, could those denied an intervention complain? Importantly, who would fund any intervention? And who would deliver it? Another AI?

@@

Reliance on AI can short-circuit the learning process. Statements like "in the very near future AIs will likely be the primary way we access knowledge" [@hie_how_2023] convey the idea that prompt engineering will become an essential skill; perhaps even the skill to end all skills^[Prompt engineering is a pompous name for a task that more closely resembles reading the entrails of a Large Language Model than actual engineering. It is a task that cannot possibly be performed in a systematic way, and that offers no guarantees that it will work in the same way again, especially after a model is updated. No genuine engineer would do their work in such fashion--and calling the activity of writing prompts "engineering" is a blatant attempt to steal the competence and reputation of actual engineering.].

It is important to note that, even if prompt "engineering" offered some value to students, the skill is inextricable tied to a singular technology, with little transferability potential. As @mollick_instructors_2024 demonstrate, prompt "engineering" is not even transferable _between_ different LLMs (see p. 5 and p. 36). Good skills transfer between contexts. Learning how to form an effective bibliographical query would work at retrieving relevant resources at McMaster, at the University of Toronto, and at the New York Public Library. Learning to identify relevant content would work while reading a novel, a book of history, or a text of econometrics. Writing, the slow and demanding process of converting thoughts into words, is essential to create a style of communication that is unique to each person. It might be argued, in fact, that writing _is_ learning [@klein_trends_2016].

In a sense, an AI is like a crutch. It can help someone to accomplish a task more easily, but it will not help if used constantly. However, what the public is being promised is not a crutch: it is an All-Terrain Vehicle that the end user cannot hope to fully master^[Reports of people getting their fingers crushed by Tesla's cybertruck could be a result of 1) people trusting the technology more than they should; and 2) not understanding the technology: apparently the algorithm uses increasing force to close the trunk if it fails to latch on back-to-back cycles. Hence, the trunk will not slash a banana on its first attempt to latch, but will crush a finger on the second attempt [@lambert_tesla_2024].] due to its blackbox nature--in the sense that 1) neural networks are blackboxes; and 2) the development of AIs itself is a blackbox wherein things can more easily change based on the interests of their corporate owners^[Sometimes said interest may be relatively benign, like avoiding reputational damage [by correcting a wrong](https://www.bbc.com/news/technology-65202597), but there is no guarantee that corporate interests will always align with those of the public, and in fact, they seldom will as long as there are only a handful of American companies that offer an "essential" service.] than in response to the interests of their relatively small customers^[And this refers not only to individuals whose skills become yoked to ChatGPT and MidJourney. In 2022, McMaster University operated on revenues of less than [\$843 million](https://financial-affairs.mcmaster.ca/app/uploads/2023/06/2023-24-Consolidated-Budget-FINAL.pdf) (CAD). Microsoft, in contrast, operated on revenues of approximately [\$198,000 millions](https://www.microsoft.com/investor/reports/ar22/) (USD) in 2022. To put things in perspective (and using the average exchange rate of 2022), if McMaster used _all_ of its 2022 revenue to pay Microsoft, that would leave McMaster with zero dollars and still account for less than _one third of one percent point_ of Microsoft's revenue. A mid-sized Canadian university is like an ant scuttling under the economic boot of Microsoft. In truth, Microsoft is so incredibly large, that it feels like it can stand up to even the US Government. As revealed by the case of Solar Winds, Microsoft refused to fix a known security flaw that left the US Government vulnerable to a Russian hack, and probably did so with little fear of ever facing accountability [@burke_microsoft_2024].].

@@

But let us assume for a moment that using AIs in their current form does becomes an in-demand tool in the short term. What is the expected longevity of skills like prompt "engineering"?

Apart from a few methodological developments (like transformers), improvement in the performance of AIs (say, from ChatGPT-2 to the latest version), mostly are the result of using bigger datasets. This has led to a rush, as developers try to improve their models by training them on more and more data^[A rush that feels at least as frenetic as any gold rush in the past. And, while the gold rushes of the 19th century were relatively democratic due to their low entry costs--both in terms of immigration and tools--the current data rush has enormous entry costs and is largely dominated by venture capital (that is, equity investments that support pre-launch, launch, and early stages of a business) and established large corporations.]. But, despite the large number of text items and images available in digital form, there is still only a finite amount of data to be had.

As of this writing, the few big players that can afford to grab the data, are mostly done with all the data grabbing, sometimes by morally dubious, and at times by possibly illegal means [@clark_ex-amazon_2024]. IBM, for example, uses data drawn from the Project CodeNet Dataset. The data are "sourced" from two online judge web sites, AIZU Online Judge and AtCoder [@ibm_automatic_2021]. According to IBM "[the] end goal of CodeNet is to enable developers to create systems that can modernize existing codebases, as well as fix errors and security vulnerabilities in code". When we inspect the Terms of Use of AIZU (a Japanese academic project), "Commercial uses of the registered codes are prohibited". IBM studiously avoids the word "commercial" in its posts about LLM, foundational models, and CodeNet. One might ask if IBM is doing all this research as a non-commercial effort? Will those developers creating systems, modernizing databases, fixing errors, etc., be able to access CodeNet for free, and use the systems for education and research only? What does "commercial uses" even mean?

@@

OpenAI, as a leader in data grabbing, turned to Stack Overflow for more data [@stack_overflow_stack_2024]. Stack Overflow [bills itself](https://stackoverflow.com) as "[a] community-based space to find and contribute answers to technical challenges, and one of the most popular websites in the world". Said community was [unhappy](https://favtutor.com/articles/stack-overflow-community-not-happy-openai/) (to say the least) with the way Stack Overflow dealt with OpenAI, and has come to believe that the terms were abusive: while Stack Overflow's terms of service state that posters cannot revoke permission to use, the site also operates under a Creative Commons license that requires attribution, something that AIs notoriously _do not do_. Some members of the community protested the deal with OpenAI by trying to withdraw or change their posts, in what Ars Technica unhelpfully termed [sabotage](https://arstechnica.com/information-technology/2024/05/stack-overflow-users-sabotage-their-posts-after-openai-deal/). It is an open question how long Stack Overflow will remain a community-based space, especially after it started treating its members as unwitting, and even unwilling providers of data for OpenAI.

@@

The implication should be clear: training models with the output of models is a recipe for less varied outputs. We may be witnessing _Peak ChatGPT_, and future versions of AIs, instead of improving, may begin to regress to a bland version of their own previous outputs. And this is before we even consider the possibility of AIs being deliberately "poisoned". As backlash against a data rush that is perceived as exploitative and abusive grows, new tools are being developed to generate data that are manipulated in ways that are too subtle to be detectable by a human, but that are deleterious to the training process of a GenAI [@salman_raising_2023]. One such tool is @shan_nightshade_2024's _Nightshade_, an algorithm that works to corrupt prompts used in text-to-image generation, and that notably bleeds through to poison related concepts. Another tool is @xu_shadowcast_2024's _Shadowcast_, an algorithm that poisons samples to trick an AI into misidentifying class labels, and that also can perform persuasion attacks, leveraging image-to-text generation capabilities to craft nonsensical narratives.
